Watch Out for the Bully! Job Interference Study on Dragonfly Network

### 2016-04-06 study_workload_1.tex
section 4 tweaks, took out comment blob

%Explain RM
In most cases for this study, the random placement policy behaves similarly across routing policies.
Random placement uniformly distributes MPI ranks over the network, 
balancing the resulting traffic load. 
As shown in Figures~\ref{fig:global-channel-traffic} and~\ref{fig:local-channel-traffic}, 
no router suffers exceptionally high volume of traffic on its local and global channels. 
When random placement is coupled with minimal routing (RM), less traffic is generated on account of
the packets avoiding intermediate forwarding. At the same time, there is still significant congestion on local channels due to the lack of ability of packets to traverse non-minimal routes, falling into the same trap as the contiguous-minimal configuration.
\TODO{check this}
%Explain RA, RPA
Coupled with (progressive) adaptive routing,
no router suffers exceptionally long saturation times on their global channels when random placement is in use, 
as shown in Figure~\ref{fig:global-channel-stime} and~\ref{fig:local-channel-stime}. Further, in comparison to contiguous allocations, random allocations result in a more evenly distributed load on the resulting channels, as expected.

@@

For the MultiGrid and CrystalRouter applications (Figures~\ref{fig:mg-commtime} and
\ref{fig:cr-commtime}, respectively), using random allocation with any routing
method results in performance improvements over contiguous allocations, which is
largely in agreement with previous work (see Section~\ref{sec:related work}).
\TODO{did previous work make this strong a claim?}. The high-radix nature of the
network topology ensures that the benefits from the resulting load balancing
outweigh the costs of extra hops for point-to-point messages.

@@

MultiGrid shows roughly similar behavior for contiguous allocations, but
different behavior along the local channels. There is a significant variation in
the traffic distribution of local channels, even with adaptive routing, which
should have the net effect of distributing said traffic.

AMG shows a similar level of variation to MultiGrid in this case. However, it is
the least communication-intensive application of the three by a significant
factor. In effect, the routers serving the AMG application are being utilized by
MultiGrid and CrystalRouter, resulting in AMG traffic queuing up in their
respective routers. Since the relative traffic is so low, AMG ends up
experiencing significant slowdowns due to this effect. We refer to this phenomenon
as AMG being ``bullied" by MultiGrid and CrystalRouter.

\begin{comment} JJ - some more reorg here, also a few of the observations don't match the data

The routers serving MultiGrid and CrystalRouter have a high traffic volume on both local and global channels when contiguous placement is in use, 
shown as the dashed line in Figure~\ref{fig:mg-lc-traffic},~\ref{fig:mg-gc-traffic} and Figure~\ref{fig:cr-lc-traffic},~\ref{fig:cr-gc-traffic}. 
On the other hand, the random placement can uniformly distribute the traffic of MultiGrid and CrystalRouter over the network, 
alleviate local congestion and balance the traffic load among all routers, 
shown as the solid lines in Figure~\ref{fig:mg-lc-traffic},~\ref{fig:mg-gc-traffic} and Figure~\ref{fig:cr-lc-traffic},~\ref{fig:cr-gc-traffic}. 
MultiGrid and CrystalRouter benefit from random placement by getting their high volume of traffic being redirected 
from their over-loaded routers to the less busy routers in the network. 
In this case, a majority of those less busy routers are belonging to AMG.

Since AMG is less communication-intensive, 
the routers serving AMG have small volume of traffic on local and global channels when the contiguous placement is use, 
shown as the dashed line in Figure~\ref{fig:amg-lc-traffic},~\ref{fig:amg-gc-traffic}. 
When random placement is in use, 
more routers are involved in serving AMG, 
and the traffic skyrockets on the local and global channels, 
shown as the solid lines in Figure~\ref{fig:amg-lc-traffic},~\ref{fig:amg-gc-traffic}. 
The increased traffic comes from the communication-intensive MultiGrid and CrystalRouter. 
AMG behaves abnormally when random placement is in use 
because the routers serving AMG have to take the traffic from MultiGrid and CrystalRouter. 
The extra traffic from the other communication-intensive applications 
takes advantage the network resources belonging to AMG and slows down AMG's communication. 
We refers this phenomenon as AMG being ``bullied" by MultiGrid and CrystalRouter.
\end{comment}

@@

In summary, based on the simulations presented in Section~\ref{sec: workload-1 network analysis} and \ref{sec: workload-1 app analysis}, we have made the following observations.

@@

\emph{The performance of communication-intensive jobs in the system improves using a random allocation policies}
Both CrystalRouter and MultiGrid, the two most communication-intensive jobs,
both saw improved distributions of communication performance when moving to a
random allocation. Again, this matches what is seen in the literature.

\emph{The performance of less communication-intensive jobs in workload regresses when random placement and adaptive routing are in use.} 
AMG in Workload~\Rmnum{1} are ``bullied" by its concurrently running communication-intensive peers, MultiGrid and CrystalRouter.
AMG shares routers and groups with MultiGrid and CrystalRouter under random placement. 
The traffic from MultiGrid and CrystalRouter is (re)directed to the routers that serve AMG, 
slowing down AMG's communication. 
\footnote{We have tried three different congestion ``sensing'' schemes in the literature for adaptive routing~\cite{won-prog-adaptive}. Although there are some variations between the results, none of the congestion sensing scheme prevent the ``bully'' behavior.}

>>>>

%Explain RM
In most cases for this study, the random placement policy behaves similarly across routing policies.
Random placement uniformly distributes MPI ranks over the network, 
balancing the resulting traffic load. 
As shown in Figures~\ref{fig:global-channel-traffic} and~\ref{fig:local-channel-traffic}, 
no router experiences an exceptionally high volume of traffic on its local and global channels. 
When random placement is coupled with minimal routing (RM), less traffic is generated on account of
the packets avoiding intermediate forwarding. At the same time, there is still significant congestion on local channels due to the lack of ability of packets to traverse non-minimal routes, falling into the same trap as the contiguous-minimal configuration.
\TODO{check this}
%Explain RA, RPA
Coupled with (progressive) adaptive routing,
saturation times are effectively minimized on both global and local channels when random placement is in use, 
as shown in Figure~\ref{fig:global-channel-stime} and~\ref{fig:local-channel-stime}. Further, in comparison to contiguous allocations, random allocations result in a more evenly distributed load on the resulting channels, as expected.

@@

For the MultiGrid and CrystalRouter applications (Figures~\ref{fig:mg-commtime} and
\ref{fig:cr-commtime}, respectively), using random allocation with any routing
method results in performance improvements over contiguous allocations, which is
largely in agreement with the literature (see Section~\ref{sec:related work}).
\TODO{did previous work make this strong a claim?}. The high-radix nature of the
network topology ensures that the benefits from the resulting load balancing
outweigh the costs of extra hops for point-to-point messages.

@@

MultiGrid shows roughly similar behavior for contiguous allocations, but
different behavior along the local channels. There is a significant variation in
the traffic distribution on local channels, even with adaptive routing, which
nevertheless has the net effect of reducing the maximal traffic load.

AMG shows a similar level of variation to MultiGrid in this case. However, it
is the least communication-intensive application of the three by a significant
factor. As evidenced by the wide gap between the router traffic in contiguous
and random placement configurations, the routers serving the AMG application
are being utilized by MultiGrid and CrystalRouter, resulting in AMG traffic
contending with traffic of other applications. The net effect, as shown in
Figure~\ref{fig:amg-commtime}, is significant slowdowns for AMG.
We refer to this phenomenon as AMG being ``bullied" by MultiGrid and CrystalRouter.

@@

In summary, based on the simulations presented in Section~\ref{sec: workload-1 network analysis} and \ref{sec: workload-1 app analysis}, we make the following observations.

@@

\emph{The performance of communication-intensive jobs in the system improves through use of random allocation policies.}
Both CrystalRouter and MultiGrid, the two most communication-intensive jobs,
both saw improved distributions of communication performance when moving to a
random allocation. Again, this matches what is seen in the literature.

\emph{The performance of less communication-intensive jobs in workload regresses when random placement and adaptive routing are in use.} 
AMG in Workload~\Rmnum{1} is ``bullied" by its concurrently running communication-intensive peers, MultiGrid and CrystalRouter.
AMG shares routers and groups with MultiGrid and CrystalRouter under random placement. 
The traffic from MultiGrid and CrystalRouter is (re)directed to the routers that serve AMG, 
slowing down AMG's communication. 
\footnote{We have tried three different congestion ``sensing'' schemes in the literature for adaptive routing~\cite{won-prog-adaptive}. Although there are some variations between the results, none of the congestion sensing scheme prevent the ``bully'' behavior.}
